Title here

### 2017-05-15 portfolio/text/anderson-etal-rev-pfx.tex
Editing for reviewer comments

We used combinations of state and federal fishing permits held by an individual
in a specific year to define fishing ``strategies'', representing the species
caught (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g.~a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were caught a single species and differed only in region fished. For
example, someone fishing herring roe in Kodiak and someone fishing herring roe
in Alaska Peninsula were both considered to have a ``herring roe'' permit
strategy. For salmon permits, we formed salmon strategies based on similarities in
gear type and species composition (Supporting Information).

>>>>

We used combinations of state and federal fishing permits held by an individual
in a specific year to define fishing ``strategies'', representing the species
caught (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g.~a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were caught a single species and differed only in region fished. For
example, someone fishing herring roe in Kodiak and someone fishing herring roe
in Alaska Peninsula were both considered to have a ``herring roe'' permit
strategy. For salmon permits, we formed salmon strategies based on similarities in
gear type and species composition (Supporting Information).
Gear types and permits are closely linked since in most cases a fisher would require a different permit to use different gear. The ability to diversify without changing gear or permits is therefore driven by biological availability and small-scale choices by fishers, such as fishing early versus late in a year or fishing in particular locations.
% Response to reviewer comment A5

### 2017-05-15 portfolio/text/ms.Rmd
Editing for reviewer comments

\caption{Effect of specializing on year-to-year changes in revenue and revenue
variability for individual permit holders \emph{within} the `r nstr` most common
strategies. Changing species diversity from year-to-year is associated with
higher variability for individuals within the strategies above the zero line.
Specializing is associated with greater expected revenue ratios to the right of
the zero line and lower expected revenue ratios to the left of the zero line.
For example, individuals with setnet permits for salmon in Prince William Sound,
Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula (second purple dot from
right) target primarily sockeye salmon and experience a $\sim$1.5\% increase in
their year-to-year revenue ratio (horizontal axis) and a $\sim$0.5\% increase in
their revenue variability (vertical axis) for a 1\% decrease in their ratio of
year-to-year species diversity. These effects are estimated holding fishing
season length constant from year-to-year. Dots and line segments represent
medians and 50\% credible intervals of posteriors. The effect of generalizing
(increasing species diversity) is similar (Fig. S\figmirror).}

@@

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014; @cline2017], but our
findings within strategies, with respect to individual year-to-year behavior,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. 
Similar efficiency gains from specialization are well known in 
agricultural fields [@ikerd1990; @abdulai2000].
<!-- For example, rice farmers in -->
<!-- Ghana can increase their efficiency by devoting more land and labor to rice -->
<!-- production compared to other crops [@abdulai2000]. -->
Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, management, biological, or
environmental conditions change. For example, U.S. farmers specialized on wheat,
corn, or soybeans were hit harder than diversified farmers in the early 1980s
when a financial crisis hit [@ikerd1990].
<!-- AB: We might want to include more examples / references from fisheries.
herring - affected by closures? value declined...-->
Third, diversifying---a lack of
specialization---can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource. For example, Ugandan
households relying on forestry resources for income increase their income
portfolio diversity following downward economic shocks [@debela2012].
Because specialization inherently holds risk from ``putting all one's eggs in one
basket,'' it is not suprising that a decline in the most valuable resource
would lead to a shift towards diversification.

@@

There are many possible barriers to diversification for individuals relying on
natural resources for their livelihood. A primary barrier is the cost of
diversifying. For commercial fishers, purchasing additional licenses and gear
typically represents a substantial expense. Related barriers are apparent in
other natural resource sectors. For example, diversification has been positively
linked with total income or access to financial capital for farmers
[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
those relying on forest resources [@illukpitiya2008]. Another major barrier to
diversification can be management restrictions. In commercial fisheries,
regulations that have been shown to reduce overfishing---including improved
assessments [@melnychuk2017] and individual quota systems 
[@costello2008]---may ultimately restrict individual diversification [@smith1986; @kasperski2013]
or at least make developing a diverse fishing portfolio more expensive. However,
individual quota systems, in particular, may also reduce income variability
(Fig. S\figifq A) perhaps offsetting the effects of reduced diversification
[@holland2016].

Our analysis provides a synthetic view of the effects of species diversity on 
individuals over 30 years in one of the most productive fishing regions in the 
world. However, this large temporal and geographical scope potentially obscures 
time or regional-variation in the consequences of diversification on revenue and
revenue variability. Furthermore, in addressing the overall pattern, we have
averaged out the potential signal of several major discrete events, such as the
*Exxon Valdez* oil spill in 1989, that was a significant disruption to the
management, ecological, and social structure of fisheries in Alaska
[@paine1996]. In addition to revenue or income variability, it is important to
understand the benefits of diversification in the context of rare and extreme
disturbances which occur in natural systems worldwide [@anderson2017].

With increasing environmental variability and extreme events in a changing
climate [@easterling2000], it is particularly important that individuals who
rely on natural resources as a primary source of income, such as fishers,
farmers, and forestry workers, have effective strategies to reduce risk. While
diversification can provide a powerful form of risk buffering, we have shown
that the relationship between individual changes to diversification and income
variability is complex and context specific. Furthermore, external market forces
can interact with management regulations, which have other considerable
benefits, to restrict individual diversification opportunities and promote
specialization. Although variability in individual revenue may not be actively
incorporated into management decisions, analyses such as ours help to explain
the range of diversification strategies available, trade-offs inherent in such
management decisions, and how these trade-offs may change over time.

>>>>

\caption{Effect of specializing on year-to-year changes in revenue and revenue
variability for individual permit holders \emph{within} the `r nstr` most common
strategies. Changing species diversity from year-to-year is associated with
higher variability for individuals within the strategies above the zero line.
Specializing is associated with greater expected revenue ratios to the right of
the zero line and lower expected revenue ratios to the left of the zero line.
For example, individuals with setnet permits for salmon in Prince William Sound,
Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula (second purple dot from
right) target primarily sockeye salmon and experience a $\sim$1.5\% increase in
their year-to-year revenue ratio (horizontal axis) and a $\sim$0.5\% increase in
their revenue variability (vertical axis) for a 1\% decrease in their ratio of
year-to-year species diversity. These effects are estimated holding fishing
season length constant from year-to-year. Dots and line segments represent
medians and 50\% credible intervals of posteriors. 
In other words, the axes represent coefficient values from the strategy-level slopes. Moving along an axis (say from left to right on the horizontal axis representing the effect of specializing on revenue) represents an increase in how steep the slope is between specialization and revenue.
The effect of generalizing
(increasing species diversity) is similar (Fig. S\figmirror).}

@@

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014; @cline2017], but our
findings within strategies, with respect to individual year-to-year behavior,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. 
Similar efficiency gains from specialization are well known in 
agricultural fields [@ikerd1990; @abdulai2000].
Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, management, biological, or
environmental conditions change. For example, U.S. farmers specialized on wheat,
corn, or soybeans were hit harder than diversified farmers in the early 1980s
when a financial crisis hit [@ikerd1990].
Third, diversifying---a lack of
specialization---can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource. For example, Ugandan
households relying on forestry resources for income increase their income
portfolio diversity following downward economic shocks [@debela2012].
Because specialization inherently holds risk from ``putting all one's eggs in one
basket,'' it is not suprising that a decline in the most valuable resource
would lead to a shift towards diversification.
<!-- Response to A1 -->
Conversely, a specialization may appear as a response to increased revenue. A fisher who achieves high revenue from a hotspot of a particular fish species may choose to stop or pause fishing thereby indirectly limiting catch diversity. For example, pink and chum salmon are the dominant species caught in purse seine fisheries in Prince william Sound and their run timing rarely overlaps. If a fisher achieves high revenue with one species that fisher may cut the season short.

@@

<!-- there are highly variable costs among fisheries with gear and permits in particular 
but it is not clear that costs vary systematically with the diversity of the catch 
there are opportunity costs to ... -->
<!-- gear costs incur considerable capital investment and some maintenance costs 
Permit types and gear types are closely linked and therefore fishers largely rely on the same gear regardless of whether you fish a specialized or diverse catch 
people have limited flexibility to modify gear within permits which are largely determined by regulations 
some permit combinations would allow a fisher to fish multiple permits with a single boat while others would require multiple boats
just expand our list of cost factors and call it a day 
costs such as leasing vessels from other fishers
maintaining equipment  -->

There are many possible barriers to diversification for individuals relying on
natural resources for their livelihood. A primary barrier is the cost of
diversifying. For commercial fishers, purchasing additional licenses and gear
typically represents a substantial expense. Related barriers are apparent in
other natural resource sectors. For example, diversification has been positively
linked with total income or access to financial capital for farmers
[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
those relying on forest resources [@illukpitiya2008].
<!-- Response to reviewer comment A2 -->
While we were unable to incorporate costs, it is possible that including license and gear purchase costs along with others such as costs of gear maintenance, catch transportation, and selling to a processor would unveil a different relationship between catch diversification and profits than the relationships between catch diversification and revenue shown here.
Another major barrier to
diversification can be management restrictions. In commercial fisheries,
regulations that have been shown to reduce overfishing---including improved
assessments [@melnychuk2017] and individual quota systems 
[@costello2008]---may ultimately restrict individual diversification [@smith1986; @kasperski2013]
or at least make developing a diverse fishing portfolio more expensive. However,
individual quota systems, in particular, may also reduce income variability
(Fig. S\figifq A) perhaps offsetting the effects of reduced diversification
[@holland2016].

Our analysis provides a synthetic view of the effects of species diversity on 
individuals over 30 years in one of the most productive fishing regions in the 
world. 
<!-- Response to reviewer comment B2 -->
Although we focus on Alaska, and expect our results are largely generalizable to other productive and intensely managed fishing regions, it is possible that the benefits of diversification would be stronger in regions where fish stocks are highly variable or in less intensely managed regions where permit costs are lower and more freely available.
Overall, the large temporal and geographical scope of our analysis potentially obscures 
time or regional-variation in the consequences of diversification on revenue and
revenue variability. Furthermore, in addressing the overall pattern, we have
averaged out the potential signal of several major discrete events, such as the
*Exxon Valdez* oil spill in 1989, that was a significant disruption to the
management, ecological, and social structure of fisheries in Alaska
[@paine1996]. In addition to revenue or income variability, it is important to
understand the benefits of diversification in the context of rare and extreme
disturbances which occur in natural systems worldwide [@anderson2017].

With increasing environmental variability and extreme events in a changing
climate [@easterling2000], it is particularly important that individuals who
rely on natural resources as a primary source of income, such as fishers,
farmers, and forestry workers, have effective strategies to reduce risk. While
diversification can provide a powerful form of risk buffering, we have shown
that the relationship between individual changes to diversification and income
variability is complex and context specific. Furthermore, external market forces
can interact with management regulations, which have other considerable
benefits, to restrict individual diversification opportunities and promote
specialization. 
From a management perspective, metrics such as interannual variability in individual fishing revenue or fraction of single permit holders within communities could serve as bioeconomic indicators.
Although variability in individual revenue is not actively
incorporated into management decisions, analyses such as ours help to explain
the range of diversification strategies available, trade-offs inherent in such
management decisions, and how these trade-offs may change over time.
